\Title

### 2023-01-19 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

The response of a biological neuron depends on the precise timing of afferent spikes. This temporal aspect of the neuronal code is essential in understanding information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras. However, most artificial neuronal models do not take advantage of this minute temporal dimension. Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous delays which we apply to the computer vision task of motion detection. Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree. We show this can be formalized as a time-invariant logistic regression which can be trained using labeled data. We apply this model to solve the specific computer vision problem of motion detection, and demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras. In particular, we quantify how the accuracy of the model can vary with the total computational load. This end-to-end event-driven computational brick could help improve the performance of future Spiking Neural Network (SNN) algorithms and their prospective use in neuromorphic chips.

@@

Let us now formally define the Heterogeneous Delays Spiking Neural Network (HD-SNN) model, as well as the task and how we will learn to solve it. First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays. Our HD-SNN model will be tested on a computationally relevant task of detecting motion detection from an event-based sensory input. By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory. Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina. The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays. To this end, we will extend the HD-SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels. We will finally describe its implementation on conventional computers. 

@@

In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} (see Figure~\ref{fig:model}-(a)). This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events. Events are typically ordered by their time of occurrence. Each event has a time of occurrence $\timev_\arank$  and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras). This defines an address space $\presynaddrspace$ which consists of the set of possible addresses. In a neurobiological recording, this can be the identified set of neurons. For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras).

>>>>

The response of a biological neuron depends on the precise timing of afferent spikes. This temporal aspect of the neuronal code is essential in understanding information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras. However, most artificial neuronal models do not take advantage of this minute temporal dimension. Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays which we apply to the computer vision task of motion detection. Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree. We show this can be formalized as a time-invariant logistic regression which can be trained using labeled data. We apply this model to solve the specific computer vision problem of motion detection, and demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras. In particular, we quantify how the accuracy of the model can vary with the total computational load. This end-to-end event-driven computational brick could help improve the performance of future Spiking Neural Network (SNN) algorithms and their prospective use in neuromorphic chips.

@@

Let us now formally define the Heterogeneous Delays Spiking Neural Network (HD-SNN) model, as well as the task to solve and how we will learn to solve it. First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays. Our HD-SNN model will be tested on a computationally relevant task of detecting motion detection from an event-based sensory input. By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory. Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina. The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays. To this end, we will extend the HD-SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels. We will finally describe its implementation on conventional computers. 

@@

In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} or \emph{events} (see Figure~\ref{fig:model}-(a)). This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events. Events are typically ordered by their time of occurrence. Each event has a time of occurrence $\timev_\arank$  and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras). This defines an address space $\presynaddrspace$ which consists of the set of possible addresses. In a neurobiological recording, this can be the identified set of neurons. For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras).
